Towards Proof-Reconstruction of Problems in Classical Propositional Logic

### 2017-10-17 paper/sections/proof-reconstruction.tex
[ paper ] minor changes.

Now, we have removed redundancies in the disjunctions by applying the
$rm_{∨}$, we can define a similar function to work with conjunctions.

@@

The \canonicalize rule is an overloaded inference rule that performs
mainly normalization of the proposition input by using theorems
described in the section above. Metis can introduce axioms, and definitions
by using it. This rule can convert a \texttt{fof} %TODO
formula in clausal form into a CNF clause. Since the nature of this rule
consists of dealing with CNF clauses, to justify its reasoning, our strategy
mainly consists of checking the equality of normalized negative form between
the source and the target formula. If it fails, we try to reorder the
conjunctive normal form of the source formula to match
the conjunctive normal form of the target formula.

@@

\begin{proof}\hspace{3mm}
\begin{itemize}
\item If $φ ≡ ψ$ by substitution theorem we conclude $Γ ⊢ ψ$.
\item If $ψ ≡ form(nnf(φ))$,
\begin{equation*}
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ φ$}
    \RightLabel{lem-nnf}
    \UnaryInfC{$Γ ⊢₂~nnf(φ)$}
    \RightLabel{lem-form}
    \UnaryInfC{$Γ ⊢ form(nnf(φ))$}
    \AxiomC{$ψ ≡ form(nnf(φ))$}
    \RightLabel{subst.}
    \BinaryInfC{$Γ ⊢ ψ$}
  \end{bprooftree}
\end{equation*}
\item If $γ := cnf(ψ) ≡ reorder_{∧∨}(cnf(φ), cnf(ψ))$,
\begin{equation*}
  \begin{bprooftree}
     \AxiomC{$Γ ⊢ φ$}
     \RightLabel{thm-cnf}
     \UnaryInfC{$Γ ⊢ cnf(φ)$}
     \RightLabel{thm-reorder$_{∧∨}$}
     \UnaryInfC{$Γ ⊢ reorder_{∧∨}(cnf(φ), cnf(ψ))$}
     \AxiomC{$γ$}
    \RightLabel{subst}
    \BinaryInfC{$Γ ⊢ cnf(ψ)$}
    \RightLabel{thm-inv-cnf.}
    \UnaryInfC{$Γ ⊢ ψ$}
  \end{bprooftree}
\end{equation*}
\end{itemize}
\end{proof}

@@

The \clausify rule transforms a
propositional formula in its clausal normal form, a conjunction
of clauses. Where a \emph{clause} is the disjunction of zero or more
literals and a \emph{literal} is an atom (positive literal) or a
negation of an atom (negative literal). Be aware this kind of conversion between one formula to its clausal normal form is not unique, and \Metis has a customized approach to perform that transformation.
In the following \Metis \TSTP derivation, we see a clear example, where
clausify expands the formula \texttt{normalize\_0\_0} by using distributive
laws.

>>>>

Now, we have removed redundancies in the disjunctions by applying
$rm_{∨}$ function, we can define a similar function to work with conjunctions.

@@

The \canonicalize rule is an overloaded inference rule that performs
mainly normalization of the proposition input by using theorems
described in the section above.  It can introduce axioms and definitions
in the problem. This rule convert a \texttt{fof} %TODO
formula in clausal form to a \CNF clause but also it perform some
simplification in the process (removing ⊤ and ⊥ from the formula).

Since this rule mostly consists of dealing with \CNF clauses, to justify its reasoning, our strategy mainly consists of checking the equality of normalized negative form between the source and the target formula. If it fails, we try to reorder the conjunctive normal form of the source formula to match with
the conjunctive normal form of the target formula.

@@

\begin{proof}\hspace{3mm}
\begin{itemize}
\item If $φ ≡ ψ$ by substitution theorem we conclude $Γ ⊢ ψ$.
\item If $ψ ≡ form(nnf(φ))$,
\begin{equation*}
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ φ$}
    \RightLabel{lem-nnf}
    \UnaryInfC{$Γ ⊢₂~nnf(φ)$}
    \RightLabel{lem-form}
    \UnaryInfC{$Γ ⊢ form(nnf(φ))$}
    \AxiomC{$ψ ≡ form(nnf(φ))$}
    \RightLabel{subst.}
    \BinaryInfC{$Γ ⊢ ψ$}
  \end{bprooftree}
\end{equation*}
\item If $γ = cnf(ψ) ≡ reorder_{∧∨}(cnf(φ), cnf(ψ))$,
\begin{equation*}
  \begin{bprooftree}
     \AxiomC{$Γ ⊢ φ$}
     \RightLabel{thm-cnf}
     \UnaryInfC{$Γ ⊢ cnf(φ)$}
     \RightLabel{thm-reorder$_{∧∨}$}
     \UnaryInfC{$Γ ⊢ reorder_{∧∨}(cnf(φ), cnf(ψ))$}
     \AxiomC{$γ$}
    \RightLabel{subst}
    \BinaryInfC{$Γ ⊢ cnf(ψ)$}
    \RightLabel{thm-inv-cnf.}
    \UnaryInfC{$Γ ⊢ ψ$}
  \end{bprooftree}
\end{equation*}
\end{itemize}
\end{proof}

@@

The \clausify rule transforms a
propositional formula in its clausal normal form, a conjunction
of clauses. Where a \emph{clause} is the disjunction of zero or more
literals and a \emph{literal} is an atom (positive literal) or a
negation of an atom (negative literal). Be aware this kind of conversion
between one formula to its clausal normal form is not unique, and \Metis has a
customized approach to perform that transformation. That is the reason we
perform some reordering for the CNF of the source formula to fulfill this
issue.

In the following \Metis \TSTP derivation, we see a clear example where
\clausify expands the formula \texttt{normalize\_0\_0} by using distributive
laws presented in conjunctive normal form conversion.
